\section{Introduction}
The proliferation of massive open online courses (MOOCs) has resulted in a
profound impact on education. As more and more learners turn to MOOCs to
educate themselves on various topics, more and more behavioral data is
being collected as part of the system on which the MOOC is offered. The
data present in these logs has the power to aid us in understanding the
behavior of students who take our MOOCs, which is mostly undetectable for
instructors of these MOOCs today due to its vast scale. As a result, the
rich data available through these MOOC logs is highly underutilized today.

What stands in the way? Clearly, intelligent systems to create concise and
digestible summaries of the massive amount of interaction data collected
are needed in order to empower the instructors of these courses. If we can
understand how users are interacting with our MOOCs, we are much more
likely to be able to make changes to these courses that positively impact
learners. While we can easily observe the changes in behavior of students
in real classrooms, MOOCs present a challenge due to their hands-off
nature and sometimes irregular schedule due to being a full-time worker.   
We view this paper as attempting to bridge this gap. Specifically,
in this paper we propose unsupervised learning methods for automatically
discovering and characterizing student learning behavior patterns or profiles from large
collections of click logs associated with MOOCs. 

We view our model as a component of a system that enables collaboration
between the machine and a human instructor to extract knowledge from these
large collections of MOOC data. Automatically extracting interpretable
patterns from the clickstream data associated with MOOCs is a necessary
step in order for instructors to identify the hidden knowledge in these
massive interaction datasets. Without the availability of a suitable model
for identifying these behavioral patterns, instructors are not empowered to
use this available data to improve their courses without expending
extraordinary amounts of manual effort.

Our proposed model is motivated by the following observations:
\begin{enumerate}
  \item Student behavior is complicated and cannot necessarily be captured
      sufficiently by rule-based methods such as those explored by
      \citet{Kizilcec:2013:LAK} and \citet{Davis:2016:EDM}. We instead
      propose to treat student behavior patterns as being characterized
      (represented) via a sequence of \emph{latent states}. This allows us
      to automatically capture patterns that we might not have been able to
      articulate clearly a priori via a series of rules, and also allows us
      to model the inherent uncertainty in assigning a student's behavior
      to a particular pattern or group.
  \item Student behavior can vary over time. Previous models that treat students
      as exhibiting only one behavioral pattern over
      time~\cite{Faucon:2016:EDM} miss out on the opportunity to understand
      student behavior dynamics in a course. We propose a latent space
      model with {\em latent state transitions} to flexibly model the
      dynamics.
  \item Analysis of student behavior can and should be performed at varying
      levels of granularity. This requires us to aggregate data over time
      with \emph{different levels of resolution}; existing models tend to come
      with a particular assumption about the resolution of time they
      consider~\cite{Faucon:2016:EDM, Kizilcec:2013:LAK, Shih:2010:EDM}. We
      propose a model that is agnostic to the time resolution considered,
      allowing it to be applied at different levels of resolution more
      naturally.
\end{enumerate}

Thus, what we propose is a \emph{latent variable approach} to mining student behavior
patterns that is \emph{probabilistic} for inference and 
does \emph{not} force assumptions about time resolutions, making it
\emph{flexible to model state changes over different time resolutions} more
easily.  More specifically, we propose a novel two-layer hidden Markov
model (2L-HMM) to discover latent student behavior patterns via
unsupervised learning on large collections of student behavior observation
sequences.  Evaluation results on a MOOC data set on Coursera demonstrate
that the 2L-HMM can effectively discover a variety of interesting interpretable
student behavior patterns at different levels of resolution, many of which
are beyond what existing approaches can discover. We show that the patterns
uncovered by the 2L-HMM capture meaningful behavior by quantitatively
showing that features extracted from a trained 2L-HMM correlate with
learning outcomes. Since our proposed methods are
unsupervised, they can potentially be applied to any MOOC data without
requiring manual annotation effort at the level of sequences. Instead,
instructors are empowered to use the latent patterns that the 2L-HMM can
discover from raw data to extract knowledge about the behaviors his/her
students exhibit in the MOOC.
